Presentation Title

### 2017-07-12 thesis/chapters/ch3-model.tex
Refactor experimental results

The goal of this chapter is to detail our system specification.
We begin our discussion with an intuitive overview of the architecture in \Cref{sec:system-overview}.
Next, we give the formal description, starting with the model and assumptions in \Cref{sec:model-assumptions}.
Then, the three protocols which make up the complete system,
namely consensus protocol (\Cref{sec:cons-protocol}), transaction protocol (\Cref{sec:tx-protocol}) and validation protocol (\Cref{sec:vd-protocol}).
Finally, some optimisations are described in \Cref{sec:optimisations}.

@@

\subsection{Relationship with $\alpha$-Synchroniser}

@@

\section{Optimisations}
\label{sec:optimisations}

Up to this point,
we have discussed our protocol in the context of the model and assumptions defined in \Cref{sec:model-assumptions}.
In this section, 
we explore a few optimisations which we can make if we relax our model.
These relaxations do not have the same theoretical elegance as our original model.
But they enable better performance and allow us to apply our design in the fully permissionless setting.

@@

>>>>

The goal of this chapter is to detail our system specification.
We begin our discussion with an intuitive overview of the architecture in \Cref{sec:system-overview}.
Next, we give the formal description, starting with the model and assumptions in \Cref{sec:model-assumptions}.
Then, the three protocols which make up the complete system,
namely consensus protocol (\Cref{sec:cons-protocol}), transaction protocol (\Cref{sec:tx-protocol}) and validation protocol (\Cref{sec:vd-protocol}).
Finally, we discuss a few variations of our main design and their respective tradeoffs in \Cref{sec:tradeoffs}.

@@

\subsection{Relationship with $\alpha$-Synchroniser}
TODO

@@

\section{Design variations and tradeoffs}
\label{sec:tradeoffs}

Up to this point,
we have discussed our protocol in the context of the model and assumptions defined in \Cref{sec:model-assumptions}.
In this section, 
we explore a few design variations which we can make if we relax our model.
These variations do not have the same theoretical elegance as our original model.
But they enable better performance, allow us to apply our design in the fully permissionless setting and improves privacy.

@@

\subsection{Global probabilistic fork detection}
TODO

### 2017-07-12 thesis/chapters/ch4-analysis.tex
Refactor experimental results

\section{Performance}
This section aims to analytically answer our research question.
That is, does the global throughput increase linearly with respect to the population size?
We begin by looking at the communication and time complexity of the consensus protocol,
and then the bandwidth requirement for a single transaction.
We build on top of those results to analyse do the global throughput analysis.

\subsection{Communication complexity of the consensus protocol}
\label{sec:acs-complexity}
The consensus protocol can bee seen as three parts.
Thus the communication complexity is the sum of these three parts.
The first part is when every node sends their CP block to all the facilitators, which is $O(Nn)$ since there are $N$ nodes and $n$ facilitators.
Or simply $O(N)$ if we consider $n$ as a constant.

@@

\subsection{Time complexity of the consensus protocol}
\label{sec:acs-time-complexity}
In order to make arguments on bandwidth or throughput, which are concepts that depend on time,
we must make some assumptions on the time complexity of consensus protocol.
The time complexity we use here is not the same as what is typically used in distributed systems.
In analysis of distributed systems, time complexity is often in terms of the number of rounds.
For example, ACS runs in a constant number of rounds because its subprotocols---reliable broadcast and binary Byzantine consensus---also run in a constant number of rounds.
However, in practice, making a unit of communication always has some overhead associated with it, for example serialising and writing it to some network socket.
Hence, for the remainder of our performance analysis, we add the following to our computational model.
For every unit of communication, we assume they take some non-negligible but constant time to perform.
Hence, from~\Cref{sec:acs-complexity}, it directly follows that the consensus protocol has a time complexity of $O(N)$.

\subsection{Communication Cost for Transactions}
\label{sec:communication-cost-for-tx}
With the analysis on the time complexity of consensus protocol,
we are ready to analyse the amount of data required to be transmitted over a link per transaction,
which we call the communication cost per transaction.
% First, we define bandwidth as the ``amount of data a link or network path can deliver per unit of
% time''~\cite{prasad2003bandwidth}.
We know that to create and then validate a transaction,
the communication cost per transaction is of $O(l)$,
where $l$ is the length of the agreed fragment.
This can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
which contains the agreed fragment.
The other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second.
Then
$$l = r_{\text{tx}} D_{\text{c}},$$
where $D_{\text{c}}$ is the duration a round of the consensus protocol.
But from \Cref{sec:acs-time-complexity},
we know that $D_{\text{c}}$ is of $O(N)$, thus the communication cost per transaction is $O(N)$.
This is intuitive because round duration would be longer if there are more CP blocks (more $N$), which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
The behaviour is also verified experimentally in \Cref{ch:implementation}.

>>>>

\section{Performance}
\label{sec:performance-analysis}
This section aims to analytically answer our research question.
That is, does the global throughput increase linearly with respect to the population size?
We begin by looking at the communication and time complexity of the consensus protocol,
and then the bandwidth requirement for a single transaction.
We build on top of those results to analyse do the global throughput analysis.

\subsection{Communication complexity of the consensus protocol}
\label{sec:cons-complexity}
The consensus protocol can bee seen as three parts.
Thus the communication complexity is the sum of these three parts.
The first part is when every node sends their CP block to all the facilitators, which is $O(Nn)$ since there are $N$ nodes and $n$ facilitators.
Or simply $O(N)$ if we consider $n$ as a constant.

@@

\subsection{Duration of the consensus protocol}
\label{sec:cons-duration}
In order to make arguments on bandwidth or throughput, which are concepts that depend on time,
we must make some assumptions regarding our computational model to make arguments on the duration of the consensus protocol.
Note that it is not the same as the time complexity typically used in distributed systems.
In analysis of distributed systems, time complexity is often in terms of the number of rounds.
For example, ACS runs in a constant number of rounds because its subprotocols---reliable broadcast and binary Byzantine consensus---also run in a constant number of rounds.
However, in practice, making a unit of communication always has some overhead associated with it, for example serialising and writing it to some network socket.
Hence, for the remainder of our performance analysis, we add the following to our computational model.
For every unit of communication, we assume they take some non-negligible but constant time to perform.
Hence, from~\Cref{sec:cons-complexity} and the fact that the consensus protocol uses a constant number of rounds,
it follows that the consensus protocol has a duration of $O(N)$.

\subsection{Communication Cost for Transactions}
\label{sec:communication-cost-for-tx}
With the analysis on the duration of the consensus protocol,
we are ready to analyse the amount of data required to be transmitted over a link per transaction,
which we call the communication cost per transaction.
% First, we define bandwidth as the ``amount of data a link or network path can deliver per unit of
% time''~\cite{prasad2003bandwidth}.
We know that to create and then validate a transaction,
the communication cost per transaction is of $O(l)$,
where $l$ is the length of the agreed fragment.
This can be seen from the fact that the largest message by far is the \texttt{vd\_resp} message,
which contains the agreed fragment.
The other messages (\texttt{tx\_req}, \texttt{tx\_resp} and \texttt{vd\_req}) are constant factors.
If we assume that every node performs transactions at a constant rate of $r_{\text{tx}}$ per second.
Then
$$l = r_{\text{tx}} D_{\text{c}},$$
where $D_{\text{c}}$ is the duration a round of the consensus protocol.
But from \Cref{sec:cons-duration},
we know that $D_{\text{c}}$ is of $O(N)$, thus the communication cost per transaction is $O(N)$.
This is intuitive because round duration would be longer if there are more CP blocks (more $N$), which means that the agreed fragments are longer (assuming nodes transact at a constant rate).
The behaviour is also verified experimentally in \Cref{ch:implementation}.

### 2017-07-12 thesis/chapters/ch5-implementation.tex
Refactor experimental results

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=1.0\textwidth]{neighbour-fixed/round-communication-cost-vs-population}
      \caption{Transactions are with fixed neighbours}
      \label{fig:consensus-comms-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=1.0\textwidth]{neighbour-random/round-communication-cost-vs-population}
      \caption{Transactions are with random neighbours}
      \label{fig:consensus-comms-random}
    \end{subfigure}%
  }
\end{figure}

\Cref{fig:consensus-comms} show the relationship between communication cost and population.
These most important aspect is that these results show a linear increase.
This reinforces our analytical result in~\Cref{sec:acs-complexity}.
Further, note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
the magnitude of the communication cost are similar.
Both peak at about $1.4 \times 10^{10}$ bytes.
This is expected because the consensus protocol is decoupled from the transaction protocol.

We are also interested in how communication costs translate to time.
Hence, for the same experiment we use duration in seconds on the vertical axis instead of communication cost and plot the~\Cref{fig:consensus-duration}
These show a similar trend as the communication cost graphs.
Hence our assumption (from~\Cref{sec:acs-time-complexity}) that for every unit of communication there is some fixed overhead is valid.

@@

\begin{figure}[h]
  \centering
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
    \caption{Transactions are with fixed neighbours}
    \label{fig:tx-comms-fixed}
  \end{subfigure}

  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
    \caption{Transactions are with random neighbours}
    \label{fig:tx-comms-random}
  \end{subfigure}
  \caption{Communication cost per verified transaction}
  \label{fig:tx-comms}
\end{figure}

Recall that in~\Cref{sec:communication-cost-for-tx} we argued that the communication cost per verified transaction is of $O(N)$.
To verify the argument, we plot the relationship between the communication cost and population size in~\Cref{fig:tx-comms}.
Again, we observe a linear relationship, which verified the argument.

However, note that there is a large difference in communication cost between the two modes of transaction.
When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
This is because agreed fragments are cached.
On the other hand,
if every node is transacting with a random node, then it is likely the case that one agreed fragment needs to be communicated for every transaction.
Hence the communication cost we see in~\Cref{fig:tx-comms-random} is much higher than in~\Cref{fig:tx-comms-fixed}.

It is interesting to note that some fluctuations exist in the graphs.
This is likely due to the way how the communication cost are recorded.
We maintain the total amount of data sent and received in a few variables.
On every message that is sent or received, we increment the relevent variable by the size of the message.
However, it is infeasible to log the values every time a message is sent or received as there is simply too much data for our build server to hold.
Therefore we only log every 5 seconds.
This does not always align with the time when a round finishes, thus some of the data may not represent reality accurately.
%  TODO this can be fixed

\section{Global Throughput}
The global throughput results are shown in~\Cref{fig:global-throughput}.
Clearly, the throughput has a linear relationship with the population size.
This result is a strong indication of the horizontal scalability which we aimed to achieve.
It also verified our analytical result in~\Cref{sec:global-throughput}.

More interestingly, is the magnitude in throughput, 
and the difference in magnitude between the two modes of transaction.
Firstly, recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have nearly 9000 transactions per second for 1200 nodes (which is about 7.5 transactions per second)?
This is due to the way validated transactions are calculated.
Transactions are between two parties, hence if every node makes two transactions per seoncd,
every node also expects to receive two transactions per second.
Hence, for every node, the TX blocks are created at 4 per second.
Every TX block must be validated.
So everyone sends 4 validation requests per second and also receives 4 per second.
Thus during one second 8 validations are performed.
The difference in magnitude is caused by the caching mechanism mentioned earlier.
If an agreed fragment needs to be transmitted to validating every transaction then it puts a toll on the network infrastructure.
The low transaction rate in~\Cref{fig:global-throughput-random} is caused by fact that the network infrastructure cannot keep up with our demand.
We demonstrate this issue from a different angle in the next section.

\begin{figure}[h]
  \centering
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/throughput-vs-population}
    \label{fig:global-throughput-fixed}
  \end{subfigure}

  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/throughput-vs-population}
    \label{fig:global-throughput-random}
  \end{subfigure}
  \caption{Global throughput}
  \label{fig:global-throughput}
\end{figure}

\section{Backlog}

TODO

\begin{figure}[h]
  \centering
  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
  \end{subfigure}

  \begin{subfigure}{\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
  \end{subfigure}
\end{figure}

>>>>

The remainder of this chapter follows the same structure as the performance analysis in~\Cref{sec:performance-analysis},
where we check our theoretical evaluation with our experimental results.

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-fixed/round-communication-cost-vs-population}
      \caption{Transactions are with fixed neighbours}
      \label{fig:round-comms-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-random/round-communication-cost-vs-population}
      \caption{Transactions are with random neighbours}
      \label{fig:round-comms-random}
    \end{subfigure}%
  }
  \caption{}
  \label{fig:round-comms}
\end{figure}

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
      \caption{Transactions are with fixed neighbours}
      \label{fig:round-duration-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
      \caption{Transactions are with random neighbours}
      \label{fig:round-duration-random}
    \end{subfigure}%
  }
  \caption{}
  \label{fig:round-duration}
\end{figure}

\Cref{fig:round-comms} show the relationship between communication cost and population.
The most important aspect is that these results show a linear increase.
This reinforces our analytical result in~\Cref{sec:cons-complexity}.
Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
the magnitude of the communication cost are similar.
Both peak at about 100 MB.
This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
Finally, the rate for which the communication cost increases is higher when the number of facilitators are higher.
This is also expected because the ACS algorithm has a $n^2$ term in its communication complexity.

We are also interested in how communication costs translate to time.
Hence, for the same experiment we also record the duration in seconds, which is shown on the vertical axis instead of communication cost and plot as before (\Cref{fig:consensus-duration}).
Interestingly, the duration is not entirely linear.
We attribute this behaviour to the extra time needed to hash the CP blocks in the consensus result to compute the luck value.
Since if $N$ increases, every node must also perform more hash operations.
These result do not conform to analytical result in~\Cref{sec:cons-duration}.
Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
For example, the luck value can be computed by the facilitators and are sent with the consensus result.
Then the non-facilitator nodes simply checks that they are correct.

@@

\begin{figure}[h]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
    \caption{Transactions are with fixed neighbours}
    \label{fig:tx-comms-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
    \caption{Transactions are with random neighbours}
    \label{fig:tx-comms-random}
  \end{subfigure}
  }
  \caption{Communication cost per verified transaction}
  \label{fig:tx-comms}
\end{figure}

Recall that in~\Cref{sec:communication-cost-for-tx} we argued that the communication cost per verified transaction is of $O(N)$.
To verify the argument, we plot the relationship between the communication cost and population size in~\Cref{fig:tx-comms}.
We observe a nearly linear relationship, which is due to the nearly linear relationship of the communication duration mentioned before.
Again, we believe the difference is minor and it is possible to remove the extra overhead.

More interestingly, there is a large difference in communication cost between the two modes of transaction.
When transacting with only one neighbour, the communication cost is low because only one agreed fragment needs to be communicated for every round in order to validate all transactions of that round.
This is because agreed fragments are cached.
On the other hand,
if every node is transacting with a random node, then it is likely the case that one agreed fragment needs to be communicated for every transaction.
Hence the communication cost we see in~\Cref{fig:tx-comms-random} is much higher than in~\Cref{fig:tx-comms-fixed}.

Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, this is due to our caching mechanism.
We send validation requests at the same rate as transactions.
Upon receiving an (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
rather than just the transaction in the original validation request.
However, it may be the case that the agreed fragments arrive later than the validation request interval.
Then it is possible to have sent two or more validation requests for the transaction in the same local agreed fragment.
In this case, the remote would respond with two or more of the same agreed fragments,
which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
The result in~\Cref{fig:tx-comms-random} reinforces our argument because it is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
% It is interesting to note that some fluctuations exist in the graphs.
% This is likely due to the way how the communication cost are recorded.
% We maintain the total amount of data sent and received in a few variables.
% On every message that is sent or received, we increment the relevent variable by the size of the message.
% However, it is infeasible to log the values every time a message is sent or received as there is simply too much data for our build server to hold.
% Therefore we only log every 5 seconds.
% This does not always align with the time when a round finishes, thus some of the data may not represent reality accurately.
%  TODO this can be fixed

\section{Global Throughput}
\begin{figure}[h]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/throughput-vs-population}
    \caption{}
    \label{fig:global-throughput-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/throughput-vs-population}
    \caption{}
    \label{fig:global-throughput-random}
  \end{subfigure}
  }
  \caption{Global throughput}
  \label{fig:global-throughput}
\end{figure}

The global throughput results are shown in~\Cref{fig:global-throughput}.
Evidently, the throughput has a linear relationship with the population size.
This result is a strong indication of the horizontal scalability which we aimed to achieve.
It also matches with our analytical result.

Note that the throughput decreases as the number of facilitators increases.
This is due to the additional communication cost for running ACS with a high number of facilitators.
That is, if the network is congested then the nodes may not have enough bandwidth to send validation responses timely.

For~\Cref{fig:global-throughput-fixed},
the magnitude of our throughput may not be self-evident at first glance.
Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (4 TX/s)?
This is due to the way validated transactions are calculated.
Transactions are between two parties, hence if every node makes two transactions per seoncd,
every node also expects to receive two transactions per second.
Hence, for every node, the TX blocks are created at 4 per second.
Validation requests are sent at the same rate, which explains the magnitude.

The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is also interesting.
It is caused by the caching mechanism mentioned earlier.
If an agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
The low transaction rate in~\Cref{fig:global-throughput-random} is caused by fact that the network infrastructure cannot keep up with our demand.
We demonstrate this issue from a different perspective in~\Cref{fig:backlog}.
The graph is plotted by counting the number of transactions and the number of validated transactions every 5 seconds for one node running in a network of 1200 nodes and 32 facilitators.
In~\Cref{fig:backlog-fixed}, the number of validated transaction changes as a step function.
This means that transactions are validated in bursts, and the validation protocol can ``keep up'' with the transactions.
Note that the horizontal ``lines'' where no new validated transactions are made are roughly 70 seconds, this matches with our result in~\Cref{fig:round-duration-fixed}.
On the other hand in~\Cref{fig:backlog-random}, the validation protocol clearly cannot ``keep up'' with the rate which the transactions are made.
As a result, the global throughput is lower when transacting with random nodes than only with neighbours.

\begin{figure}[h]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
    \caption{}
    \label{fig:backlog-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
    \caption{}
    \label{fig:backlog-random}
  \end{subfigure}
  }
  \caption{}
  \label{fig:backlog}
\end{figure}
